Supplemental Materials

### 2021-07-18 paper/paper.Rmd
updated references

One remarkable feature of human psychology is our ability to form rich mental representations of what is true, how things work, and how best to navigate the world around us. Much of this type of thinking can be understood as deriving from people’s *intuitive theories*---explanatory frameworks, analogous to scientific theories, that capture the causal and logical structure of the world within a domain. From early in life, people are driven to seek out such causal explanations and to use their mental representations of these causal systems to explain, predict, and intervene in the world; and to make sense of new information and experiences. (See Carey, 2009; Gelman & Legare, 2011; Gerstenberg & Tenenbaum; Gopnik & Wellman, 1994; Keil, 1994; Murphy & Medin, 1985; and Wellman & Gelman, 1992; for influential  treatments of this topic.)  Intuitive theories inform reasoning in a wide variety of domains, from reasoning about physical systems (e.g., Baillargeon, 2004; Battaglia, Hamrick, & Tenenbaum, 2013; Spelke, 1994; Ullman et al., 2017), to reasoning about biological processes (e.g., Carey, 1985, 2009; Medin & Atran, 1999) and about the contents of other people’s minds (e.g., Gopnik & Wellman, 1992; Wellman, 2015; Baker, Jara-Ettinger, Saxe, & Tenenbaum).

@@

Even before the coronavirus pandemic, misconceptions about vaccine safety drove public health crises around the world. Undervaccination resulted in outbreaks of measles affecting over 1000 children across the United States in 2019, and over 80,000 children across Europe in 2018 (WHO, 2018). Misconceptions about childhood vaccination---particularly the fraudulent claim that vaccines cause autism or other dangerous side-effects---have proved remarkably difficult to correct (e.g., Betsch & Sachse, 2013; Horne, Powell, Hummel, & Holyoak, 2015). Sadly, only a handful of educational interventions have been found to improve vaccine attitudes (e.g., Horne et al., 2015; Vannice et al., 2011; Wallace et al., 2006; Dube et al., 2015). This is especially  sobering in the context of the coronavirus pandemic: reluctance to be vaccinated against COVID-19 threatens our ability to end a pandemic that has already claimed millions of lives around the world.

The difficulty of convincing people to vaccinate themselves and their children illustrates a general pattern of findings showing that people often do not respond to evidence in ways that educators might wish (e.g. Lewandowsky et al., 2012). Part of the challenge in correcting misconceptions is that they are situated within larger, internally coherent belief systems---intuitive theories---that guide how people interpret and respond to evidence (e.g., Lewandowsky et al., 2012; Jern et al., 2014; Gershman, 2019).

@@

The way in which evidence affects any particular belief is unavoidably shaped by other interrelated beliefs. Duhem (1954) and Quine (1951) famously argued that scientific theories necessarily incorporate *auxiliary hypotheses* that influence how observations are interpreted: When an observation is inconsistent with a hypothesis under certain auxiliary assumptions, a scientist cannot necessarily know whether to revise the hypothesis or the auxiliary assumptions. Similarly, we argue that intuitive theories incorporate auxiliary beliefs that affect how other beliefs are revised in light of evidence (also see Gershman, 2019).

@@

We argue that correcting misconceptions requires presenting evidence in a way that is sensitive to and persuasive within the broader conceptual context in which those misconceptions reside (see Gershman, 2019; Weisman & Markman, 2017). Understanding the revision of any one specific belief (e.g., a misconception like *vaccines are dangerous*) demands a holistic understanding of the wider set of interrelated beliefs that make up people's intuitive theories.

@@

Quine argued that all of one’s beliefs are intertwined in a “web of belief” (1951; 1970)---an apt metaphor for intuitive theories. Intuitive theories consist of relations between beliefs that specify how states of affairs causally, logically, or otherwise interconnect with one another. Given this definition, we argue that any set of beliefs related by an intuitive theory can be expected to exhibit several qualitative features:

@@

Here, we sought to formalize this understanding of intuitive theories by representing an intuitive theory as a Bayesian network. Originally designed as tools for artificial intelligence (Pearl, 1988), Bayesian networks are a class of probabilistic graphical models that have been deployed widely across cognitive science (e.g. Gopnik et al., 2004; Griffiths & Tenenbaum, 2005; Jern et al., 2012). The heart of a Bayesian Network is a graphical structure ---not unlike a "web of belief"---that represents variables as nodes connected by edges, where those edges represent probabilistic or statistical relations.

For example, the graph $\text{FIRE} \rightarrow \text{SMOKE}$ represents a Bayesian network capturing the understanding  that fire causes smoke. The example of cause and effect is especially intuitive, but edges in a Bayesian network need not be causal; they are better thought of as representing “influence relations,” which may also include relationships like logical implication and set membership (Williamson, 2001). This influence relationship is represented as a probabilistic function, where the probability of smoke depends on the probability of fire.

@@

Fortunately, the formalism of Bayesian networks provides tools for navigating these challenges. Roughly, when the relationships among beliefs are unknown, we can use the qualitative features of intuitive theories to *infer* the structure of an intuitive theory from observed correlations among beliefs. Structure learning algorithms (see Scutari & Denis, 2021) provide a means for searching through the (super-exponentially large) possible space of Bayesian networks to find the structure most likely to have generated some observed data. These approaches allow us to scale the Bayesian network modelling framework beyond the simple models characteristic of early computational work on intuitive theories (e.g., Griffiths & Tenenbaum, 2005).

@@

In Study 1, we used Bayesian network structure learning algorithms to develop a cognitive model of the relationships among 14 beliefs related to vaccine hesitancy in a large sample of people. In Study 2, we replicated the success of an existing intervention emphasizing the dangers of childhood diseases (Horne et al., 2015) and used our model to provide a more detailed understanding of *how and why* this intervention increases vaccine intentions and its effects on related beliefs. In Study 3a and 3b, we drew inspiration from our model to create and validate the success of  a novel intervention aimed at debunking concerns about toxic additives. Finally, in Study 4 we tested whether people's beliefs changed coherently following relevant real-world events, by examining beliefs before and after serious outbreaks of measles in the United States in early 2019. Taken together, this work illustrates the promise of understanding belief revision in terms of intuitive theories and the utility of discovering and formalizing intuitive theories via computational models. And, it lays the foundation for empirically-validated, theory-based educational interventions that could encourage people to vaccinate their children and themselves against dangerous diseases like measles and Covid-19.

@@

These included a variety of claims about vaccines, including beliefs about (1) the overall danger of vaccines (*vaccine danger*); (2) *toxic additives in vaccines*; and (3) *vaccine effectiveness*, how effective vaccines are in preventing disease; as well as a variety of specific claims about childhood diseases like measles, mumps, and rubella, including beliefs about (4) *disease rarity* and (5) *disease severity*. Beyond this, we theorized that participants’ understanding of the infant immune system might be relevant, including beliefs that (6) the infant immune system is weak (*IIS: weakness*); (7) the infant immune system is limited in its capacity and can be easily overwhelmed (*IIS: limited capacity*); and (8) vaccines strain the infant immune system (*IIS: vaccines strain*). We also included general theories about parenting and medicine: (11) general *parental protectiveness*; (12) *parental expertise*, namely the belief that parents usually know more about their children’s health than medical experts; and (13) *medical skepticism*, including concerns about pharmaceutical companies and corruption in the medical community; and two two broad worldviews: (9) *naturalism*, a general preference for natural over artificial things; and (10) *holistic balance*, one important aspect of attitudes toward alternative medicine (e.g., “The body is essentially self-healing and the task of a health care provider is to assist the healing process”; McFadden, Hernandez, & Ito, 2010).

@@

One of the few successful educational interventions in this domain was tested by Horne et al. (2015), who found that emphasizing the dangers of preventable diseases like mumps, measles, and rubella improved participants’ attitudes toward vaccines. Their intervention included descriptions of the symptoms of mumps, measles, and rubella; vivid photos of children suffering from these diseases; and a brief story from the perspective of a parent whose child was hospitalized with measles (see SI). This “disease risk” intervention improved attitudes toward vaccines compared to an irrelevant control essay and compared to an intervention aimed at debunking concerns about vaccines causing autism.

In Study 2, we used the belief network established in Study 1 to predict how Horne et al.’s (2015) intervention would affect the beliefs in this network, including people’s intentions to vaccinate their children. Participants completed the 14 belief scales from Study 1, then returned two days later to complete either the disease risk intervention or a no-intervention control, followed by the same 14 belief scales.

Figure 3 shows the change in participants’ attitudes across all 14 belief scales, from pretest to posttest. As expected, the largest change was observed for *disease severity* beliefs---the belief directly targeted by the intervention. Most importantly, the disease risk intervention significantly increased participant’s *vaccination intentions*, conceptually replicating earlier findings regarding this intervention (Horne et al., 2015).

@@

The intervention developed by Horne et al. and employed in Study 2 is notable for being one of the few educational interventions shown to decrease vaccine hesitancy---but is disease severity the only (or most) effective target for educational interventions in this domain?

@@

More broadly, the approaches we have introduced here may shed light on long-standing theoretical questions about belief revision. Many researchers have argued that people's belief updating is subject to various biases, or is skewed by motivational factors (e.g. Kahan, 2013; Flynn et al., 2017; Kunda, 1990). Yet many of the phenomena traditionally taken as evidence for these theories can also be explained by normative models of belief revision, so long as we are willing to posit that other auxiliary beliefs can affect the relationships between a belief and evidence (e.g., Gershman, 2019; Jern et al., 2012). Adjudicating between competing views of belief revision will require us to move beyond minimal models and illustrative simulations by assessing whether people actually do hold these auxiliary beliefs, and whether their broader belief systems behave according to the rules of probabilistic inference. The Bayesian network structure learning techniques applied here are one tool we might use to build these richer models.

Though these studies were conducted before the onset of the COVID-19 pandemic and focused on childhood diseases, many similar concerns are likely to animate hesitancy toward COVID-19 vaccines, and certain insights might be readily applied to addressing them. For one, in the spirit of Horne et al.’s (2015) disease risk intervention, we suspect that education focused on how severely ill many people become and the prevalence of long-term health problems following even mild cases of COVID-19 might encourage people to accept coronavirus vaccines.

@@

Beyond this, in many countries COVID-19 and vaccines against it have been politicized in a way that other public health concerns are not. Political affiliations often imply concrete beliefs about the world that could be incorporated into a model of an intuitive theory: for instance, the belief that the coronavirus pandemic is a “hoax.” However, these affiliations are also associated with social commitments, personal values, and other identity-constituting beliefs that may be more challenging to  capture in an intuitive theory as we have conceived of it here (Van Leeuwen, 2014). Relatedly, the coronavirus pandemic has raised public awareness of many social issues that could also play an important role in people's intuitive theories of vaccination more broadly. The uneven impacts of the pandemic have laid bare inequities within the American healthcare system, and prompted a greater awareness of and reckoning with a history of racism and abuses by the medical community. Understanding of these issues, which might differ across members of majority and minoritized communities, is well worth exploration in future studies. At the same time, the concept of "herd immunity" has likely never had greater public understanding and appreciation than now. This understanding is potentially a prime target for future interventions, though one whose impacts are likely to depend on other auxiliary ethical commitments, such as moral beliefs about individuals’ obligations to their wider communities.

@@

We drew on a variety of sources, including academic articles on anti-vaccine skepticism (e.g., Dube et al., 2015; Larson et al. 2015; Salmon et al., 2015; Williams, 2014), anti-vaccine websites, and a qualitative survey with self-identified vaccine skeptics (*n*=16) via Amazon Mechanical Turk (MTurk) to generate a list of beliefs that might influence people’s decisions about whether to vaccinate their children (*vaccine intentions*): *vaccine danger*, *toxic additives in vaccines*, *vaccine effectiveness*, *disease rarity*, *disease severity, infant immune system (IIS): weakness*, *IIS: limited capacity*, *IIS: vaccines strain*, *parental protectiveness*, *parental expertise*, *medical skepticism*, *naturalism*, and *holistic balance*. (See main text for descriptions, and SI for the items included in these scales.) We translated these qualitative insights into 14 psychometrically-valid scales; each scale measured agreement with 4-6 statements, including at least one reverse-coded item, and was highly reliable (All Cronbach’s $\alpha \geq .73$). We developed 13 of these scales, and included one preexisting scale (the “holistic balance” subscale from McFadden et al., 2010).

@@

$$P(y) = 1 \big[1-(1-w_0)\prod_i(1-w_i)^{x_i}\big]\prod_j(1-w_j)^{x_j}$$ 
Here, each parent node $x$ is either a generative ($i$) or preventive ($j$) influence, that generates or prevents the child node with some probability weight ($w_i$ or $w_j$). The weight $w_0$ represents the “leak” probability, or the baseline probability that the child occurs without any other generating influences. The first term in brackets represents the probability that $y$ is generated (or, put another way, that not all of the possible generating forces fail to generate it). The second term represents the probability that it is not then prevented (i.e., that all the preventing forces fail; see Maaskant & Druzdel, 2008).

@@

In particular, we applied what we call the “generative model principle,” specifying that influence relations (edges) should flow from causes (or other generating states of affairs) to effects; instilling a core assumption in research on intuitive theories and causal explanatory reasoning (e.g. Cheng, 1997; Gopnik et al., 2004; Gerstenberg et al., 2021). We sorted the 14 measured beliefs into four tiers based on how general or abstract each belief was, reasoning that general and abstract states of affairs should be understood to cause (or otherwise generate) more specific and concrete states of affairs, so edges should flow from the general and abstract toward the specific and concrete (e.g., if natural things are better than artificial things, this should set an expectation about whether a specific class of artificial things--vaccines--is safe). We considered *holistic balance* and *naturalism* to be the most general and abstract beliefs among those we measured; followed by theories about parenting and medicine; then specific beliefs about vaccines, diseases, and the infant immune system; and finally our outcome of interest, *vaccine intentions*---the most concrete measurement of a specific intended action.

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several learning algorithms implemented in the “bnlearn” R package (Scutari, 2010) and evaluated their performance through cross validation in our training split (see SI). Following this analysis, we chose to fit our model with the score-based “hill-climbing” algorithm, scoring models via the Bayesian Information Criteria, an asymptotic estimate of the likelihood of the data under a given graph structure. This produced the network shown in Figure 1. See SI for more detailed explanations of these modeling choices.

Some caution should be exercised in interpreting the relationships in the network. Due to Bayesian networks' tight connections with structural causal networks, it is tempting to interpret the edges in this network as representing a mental causal model. But, in some cases these relationships are better thought of as probabilistic dependencies or “influence” relations (Williamson, 2001). For instance, intuitively we might imagine that vaccines having toxins could cause them to be dangerous and ineffective. These nodes are connected in the model, but the edges between them run in the opposite directions, and therefore do not support a clear causal interpretation. In other cases, our data and explicit constraints do not perfectly identify the directions of the edges. This is the case for two edges (Naturalism →  Holistic Balance and Medical Skepticism → Parental Expertise; see supplemental implementation for details). In both cases, we do not feel it is intuitive which direction the edge should flow in, nor even that it should be directed at all; we might even speculate there is an additional unmeasured variable that connects these beliefs and that would best account for their association.

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the disease risk intervention on responses to the *Vaccination Intentions* scale items, using the “brms” package for R (Buerkner, 2017). We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the disease risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention scores is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The disease risk intervention increased participants’ vaccination intention ratings (interaction effect: *b* = -0.15, 95% CI: [-0.27, -0.04]). We consider this to be a full replication of Horne, Powell, et al.’s (2015) results. See Figure 3, panel A.

@@

To model the effects of the intervention presented in Study 2 on the intuitive theory surfaced in Study 1, we augmented the belief network representing the intuitive theory with an additional node representing the information about measles, mumps, and rubella presented in the intervention. This evidence node was added as a child of the *disease severity* node, allowing the intervention to act as virtual evidence for *disease severity* (following Pearl, 1988).

@@

>>>>

One remarkable feature of human psychology is our ability to form rich mental representations of what is true, how things work, and how best to navigate the world around us. Much of this type of thinking can be understood as deriving from people’s *intuitive theories*---explanatory frameworks, analogous to scientific theories, that capture the causal and logical structure of the world within a domain. From early in life, people are driven to seek out such causal explanations and to use their mental representations of these causal systems to explain, predict, and intervene in the world; and to make sense of new information and experiences [see @carey2009; @gelman.legare2011; @gerstenberg.etal2021; @gopnik.wellman1994; @keil1994; @murphy.medin1985; @wellman.gelman1992 for influential treatments of this topic].  Intuitive theories inform reasoning in a wide variety of domains, from reasoning about physical systems [e.g., @baillargeon2004; @battaglia.etal2013; @spelke1994a; @ullman.etal2017], to reasoning about biological processes [e.g., @carey1985; @carey2009; @medin.atran1999] and about the contents of other people’s minds [e.g., @gopnik.wellman1994; @wellman2015; @baker.etal2017].

@@

Even before the coronavirus pandemic, misconceptions about vaccine safety drove public health crises around the world. Undervaccination resulted in outbreaks of measles affecting over 1000 children across the United States in 2019, and over 80,000 children across Europe in 2018 (WHO, 2018). Misconceptions about childhood vaccination---particularly the fraudulent claim that vaccines cause autism or other dangerous side-effects---have proved remarkably difficult to correct [e.g., @betsch.sachse2013; @horne.etal2015]. Sadly, only a handful of educational interventions have been found to improve vaccine attitudes [e.g., @horne.etal2015; @vannice.etal2011; @wallace.etal2006; @dube.etal2015]. This is especially  sobering in the context of the coronavirus pandemic: reluctance to be vaccinated against COVID-19 threatens our ability to end a pandemic that has already claimed millions of lives around the world.

The difficulty of convincing people to vaccinate themselves and their children illustrates a general pattern of findings showing that people often do not respond to evidence in ways that educators might wish [e.g. @lewandowsky.etal2012]. Part of the challenge in correcting misconceptions is that they are situated within larger, internally coherent belief systems---intuitive theories---that guide how people interpret and respond to evidence [e.g., @lewandowsky.etal2012; @jern.etal2014; @gershman2019].

@@

The way in which evidence affects any particular belief is unavoidably shaped by other interrelated beliefs. Duhem [-@duhem1954] and Quine [-@quine1951] famously argued that scientific theories necessarily incorporate *auxiliary hypotheses* that influence how observations are interpreted: When an observation is inconsistent with a hypothesis under certain auxiliary assumptions, a scientist cannot necessarily know whether to revise the hypothesis or the auxiliary assumptions. Similarly, we argue that intuitive theories incorporate auxiliary beliefs that affect how other beliefs are revised in light of evidence [also see @gershman2019].

@@

We argue that correcting misconceptions requires presenting evidence in a way that is sensitive to and persuasive within the broader conceptual context in which those misconceptions reside [see @gershman2019; @weisman.markman2017]. Understanding the revision of any one specific belief (e.g., a misconception like *vaccines are dangerous*) demands a holistic understanding of the wider set of interrelated beliefs that make up people's intuitive theories.

@@

Quine argued that all of one’s beliefs are intertwined in a “web of belief” [@quine1951; @quine.ullian1978]---an apt metaphor for intuitive theories. Intuitive theories consist of relations between beliefs that specify how states of affairs causally, logically, or otherwise interconnect with one another. Given this definition, we argue that any set of beliefs related by an intuitive theory can be expected to exhibit several qualitative features:

@@

Here, we sought to formalize this understanding of intuitive theories by representing an intuitive theory as a Bayesian network. Originally designed as tools for artificial intelligence [@pearl1988], Bayesian networks are a class of probabilistic graphical models that have been deployed widely across cognitive science [e.g. @jern.etal2014; @gopnik.etal2004; @griffiths.tenenbaum2005]. The heart of a Bayesian Network is a graphical structure ---not unlike a "web of belief"---that represents variables as nodes connected by edges, where those edges represent probabilistic or statistical relations.

For example, the graph $\text{FIRE} \rightarrow \text{SMOKE}$ represents a Bayesian network capturing the understanding  that fire causes smoke. The example of cause and effect is especially intuitive, but edges in a Bayesian network need not be causal; they are better thought of as representing “influence relations,” which may also include relationships like logical implication and set membership [@williamson2001]. This influence relationship is represented as a probabilistic function, where the probability of smoke depends on the probability of fire.

@@

Fortunately, the formalism of Bayesian networks provides tools for navigating these challenges. Roughly, when the relationships among beliefs are unknown, we can use the qualitative features of intuitive theories to *infer* the structure of an intuitive theory from observed correlations among beliefs. Structure learning algorithms [see @scutari.denis2021] provide a means for searching through the (super-exponentially large) possible space of Bayesian networks to find the structure most likely to have generated some observed data. These approaches allow us to scale the Bayesian network modelling framework beyond the simple models characteristic of early computational work on intuitive theories [e.g., @griffiths.tenenbaum2005].

@@

In Study 1, we used Bayesian network structure learning algorithms to develop a cognitive model of the relationships among 14 beliefs related to vaccine hesitancy in a large sample of people. In Study 2, we replicated the success of an existing intervention emphasizing the dangers of childhood diseases [@horne.etal2015] and used our model to provide a more detailed understanding of *how and why* this intervention increases vaccine intentions and its effects on related beliefs. In Study 3a and 3b, we drew inspiration from our model to create and validate the success of  a novel intervention aimed at debunking concerns about toxic additives. Finally, in Study 4 we tested whether people's beliefs changed coherently following relevant real-world events, by examining beliefs before and after serious outbreaks of measles in the United States in early 2019. Taken together, this work illustrates the promise of understanding belief revision in terms of intuitive theories and the utility of discovering and formalizing intuitive theories via computational models. And, it lays the foundation for empirically-validated, theory-based educational interventions that could encourage people to vaccinate their children and themselves against dangerous diseases like measles and Covid-19.

@@

These included a variety of claims about vaccines, including beliefs about (1) the overall danger of vaccines (*vaccine danger*); (2) *toxic additives in vaccines*; and (3) *vaccine effectiveness*, how effective vaccines are in preventing disease; as well as a variety of specific claims about childhood diseases like measles, mumps, and rubella, including beliefs about (4) *disease rarity* and (5) *disease severity*. Beyond this, we theorized that participants’ understanding of the infant immune system might be relevant, including beliefs that (6) the infant immune system is weak (*IIS: weakness*); (7) the infant immune system is limited in its capacity and can be easily overwhelmed (*IIS: limited capacity*); and (8) vaccines strain the infant immune system (*IIS: vaccines strain*). We also included general theories about parenting and medicine: (11) general *parental protectiveness*; (12) *parental expertise*, namely the belief that parents usually know more about their children’s health than medical experts; and (13) *medical skepticism*, including concerns about pharmaceutical companies and corruption in the medical community; and two two broad worldviews: (9) *naturalism*, a general preference for natural over artificial things; and (10) *holistic balance*, one important aspect of attitudes toward alternative medicine (e.g., “The body is essentially self-healing and the task of a health care provider is to assist the healing process”; [@mcfadden.etal2010]).

@@

One of the few successful educational interventions in this domain was tested by Horne et al. [-@horne.etal2015], who found that emphasizing the dangers of preventable diseases like mumps, measles, and rubella improved participants’ attitudes toward vaccines. Their intervention included descriptions of the symptoms of mumps, measles, and rubella; vivid photos of children suffering from these diseases; and a brief story from the perspective of a parent whose child was hospitalized with measles (see SI). This “disease risk” intervention improved attitudes toward vaccines compared to an irrelevant control essay and compared to an intervention aimed at debunking concerns about vaccines causing autism.

In Study 2, we used the belief network established in Study 1 to predict how Horne et al.’s [-@horne.etal2015] intervention would affect the beliefs in this network, including people’s intentions to vaccinate their children. Participants completed the 14 belief scales from Study 1, then returned two days later to complete either the disease risk intervention or a no-intervention control, followed by the same 14 belief scales.

Figure 3 shows the change in participants’ attitudes across all 14 belief scales, from pretest to posttest. As expected, the largest change was observed for *disease severity* beliefs---the belief directly targeted by the intervention. Most importantly, the disease risk intervention significantly increased participant’s *vaccination intentions*, conceptually replicating earlier findings regarding this intervention [@horne.etal2015].

@@

The intervention developed by Horne et al. [-@horne.etal2015] and employed in Study 2 is notable for being one of the few educational interventions shown to decrease vaccine hesitancy---but is disease severity the only (or most) effective target for educational interventions in this domain?

@@

More broadly, the approaches we have introduced here may shed light on long-standing theoretical questions about belief revision. Many researchers have argued that people's belief updating is subject to various biases, or is skewed by motivational factors [e.g. @kahan2013; @flynn.etal2017; @kunda1990]. Yet many of the phenomena traditionally taken as evidence for these theories can also be explained by normative models of belief revision, so long as we are willing to posit that other auxiliary beliefs can affect the relationships between a belief and evidence [e.g., @jern.etal2014; @gershman2019]. Adjudicating between competing views of belief revision will require us to move beyond minimal models and illustrative simulations by assessing whether people actually do hold these auxiliary beliefs, and whether their broader belief systems behave according to the rules of probabilistic inference. The Bayesian network structure learning techniques applied here are one tool we might use to build these richer models.

Though these studies were conducted before the onset of the COVID-19 pandemic and focused on childhood diseases, many similar concerns are likely to animate hesitancy toward COVID-19 vaccines, and certain insights might be readily applied to addressing them. For one, in the spirit of Horne et al.’s [-@horne.etal2015] disease risk intervention, we suspect that education focused on how severely ill many people become and the prevalence of long-term health problems following even mild cases of COVID-19 might encourage people to accept coronavirus vaccines.

@@

Beyond this, in many countries COVID-19 and vaccines against it have been politicized in a way that other public health concerns are not. Political affiliations often imply concrete beliefs about the world that could be incorporated into a model of an intuitive theory: for instance, the belief that the coronavirus pandemic is a “hoax.” However, these affiliations are also associated with social commitments, personal values, and other identity-constituting beliefs that may be more challenging to  capture in an intuitive theory as we have conceived of it here [@vanleeuwen2014]. Relatedly, the coronavirus pandemic has raised public awareness of many social issues that could also play an important role in people's intuitive theories of vaccination more broadly. The uneven impacts of the pandemic have laid bare inequities within the American healthcare system, and prompted a greater awareness of and reckoning with a history of racism and abuses by the medical community. Understanding of these issues, which might differ across members of majority and minoritized communities, is well worth exploration in future studies. At the same time, the concept of "herd immunity" has likely never had greater public understanding and appreciation than now. This understanding is potentially a prime target for future interventions, though one whose impacts are likely to depend on other auxiliary ethical commitments, such as moral beliefs about individuals’ obligations to their wider communities.

@@

We drew on a variety of sources, including academic articles on anti-vaccine skepticism [e.g.,  @dube.etal2015; @larson.etal2015; @salmon.etal2015; @williams2014], anti-vaccine websites, and a qualitative survey with self-identified vaccine skeptics (*n*=16) via Amazon Mechanical Turk (MTurk) to generate a list of beliefs that might influence people’s decisions about whether to vaccinate their children (*vaccine intentions*): *vaccine danger*, *toxic additives in vaccines*, *vaccine effectiveness*, *disease rarity*, *disease severity, infant immune system (IIS): weakness*, *IIS: limited capacity*, *IIS: vaccines strain*, *parental protectiveness*, *parental expertise*, *medical skepticism*, *naturalism*, and *holistic balance*. (See main text for descriptions, and SI for the items included in these scales.) We translated these qualitative insights into 14 psychometrically-valid scales; each scale measured agreement with 4-6 statements, including at least one reverse-coded item, and was highly reliable (All Cronbach’s $\alpha \geq .73$). We developed 13 of these scales, and included one preexisting scale [the “holistic balance” subscale from @mcfadden.etal2010].

@@

$$P(y) = 1 \big[1-(1-w_0)\prod_i(1-w_i)^{x_i}\big]\prod_j(1-w_j)^{x_j}$$ 
Here, each parent node $x$ is either a generative ($i$) or preventive ($j$) influence, that generates or prevents the child node with some probability weight ($w_i$ or $w_j$). The weight $w_0$ represents the “leak” probability, or the baseline probability that the child occurs without any other generating influences. The first term in brackets represents the probability that $y$ is generated (or, put another way, that not all of the possible generating forces fail to generate it). The second term represents the probability that it is not then prevented [i.e., that all the preventing forces fail; see @maaskant.druzdzel2008].

@@

In particular, we applied what we call the “generative model principle,” specifying that influence relations (edges) should flow from causes (or other generating states of affairs) to effects; instilling a core assumption in research on intuitive theories and causal explanatory reasoning [e.g. @gopnik.etal2004; @gerstenberg.etal2021; @cheng1997]. We sorted the 14 measured beliefs into four tiers based on how general or abstract each belief was, reasoning that general and abstract states of affairs should be understood to cause (or otherwise generate) more specific and concrete states of affairs, so edges should flow from the general and abstract toward the specific and concrete (e.g., if natural things are better than artificial things, this should set an expectation about whether a specific class of artificial things--vaccines--is safe). We considered *holistic balance* and *naturalism* to be the most general and abstract beliefs among those we measured; followed by theories about parenting and medicine; then specific beliefs about vaccines, diseases, and the infant immune system; and finally our outcome of interest, *vaccine intentions*---the most concrete measurement of a specific intended action.

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several learning algorithms implemented in the “bnlearn” R package [@scutari2010] and evaluated their performance through cross validation in our training split (see SI). Following this analysis, we chose to fit our model with the score-based “hill-climbing” algorithm, scoring models via the Bayesian Information Criteria, an asymptotic estimate of the likelihood of the data under a given graph structure. This produced the network shown in Figure 1. See SI for more detailed explanations of these modeling choices.

Some caution should be exercised in interpreting the relationships in the network. Due to Bayesian networks' tight connections with structural causal networks, it is tempting to interpret the edges in this network as representing a mental causal model. But, in some cases these relationships are better thought of as probabilistic dependencies or “influence” relations [@williamson2001]. For instance, intuitively we might imagine that vaccines having toxins could cause them to be dangerous and ineffective. These nodes are connected in the model, but the edges between them run in the opposite directions, and therefore do not support a clear causal interpretation. In other cases, our data and explicit constraints do not perfectly identify the directions of the edges. This is the case for two edges (Naturalism $\rightarrow$  Holistic Balance and Medical Skepticism $\rightarrow$ Parental Expertise; see supplemental implementation for details). In both cases, we do not feel it is intuitive which direction the edge should flow in, nor even that it should be directed at all; we might even speculate there is an additional unmeasured variable that connects these beliefs and that would best account for their association.

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the disease risk intervention on responses to the *Vaccination Intentions* scale items, using the “brms” package for R [@burkner2017]. We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the disease risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention scores is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The disease risk intervention increased participants’ vaccination intention ratings (interaction effect: *b* = -0.15, 95% CI: [-0.27, -0.04]). We consider this to be a full replication of Horne, Powell, et al.’s [-@horne.etal2015] results. See Figure 3, panel A.

@@

To model the effects of the intervention presented in Study 2 on the intuitive theory surfaced in Study 1, we augmented the belief network representing the intuitive theory with an additional node representing the information about measles, mumps, and rubella presented in the intervention. This evidence node was added as a child of the *disease severity* node, allowing the intervention to act as virtual evidence for *disease severity* [following @pearl1988].

@@

\newpage
# References

### 2021-07-18 supplement/supplement-main.Rmd
updated references

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package [@scutari2010a]. The first is a purely score-based “hill-climbing” algorithm  and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm, which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

>>>>

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package [@scutari2010]. The first is a purely score-based “hill-climbing” algorithm  and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm, which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.
